Mergeable Types

### 2017-02-28 paper/meta.tex
minor edit

The problem of multiple LCAs also arises in the context of source
control systems, which employ \emph{ad hoc} mechanisms to pave the way
for three-way merging.  GitHub~\cite{github}, for instance,
recursively merges LCAs to compute a virtual ancestor, which then
serves as the LCA for merging concurrent versions. This method is
demonstrated in the branching structure for a mergeable, replicated
counter as shown in Fig.~\ref{fig:criss-cross-lcas}, where LCAs 5 and
4 of 0 and 3 are merged (with their LCA being 10) to generate -1 as
the virtual LCA to merge 5 and 4. A major downside with this approach
is that it makes no guarantees on the relationship between the virtual
ancestor and its concurrent versions; the former may not even be a
legal ancestor of the latter as per the semantics of the data type.
For instance, suppose the integer type in
Fig.~\ref{fig:criss-cross-lcas} represents a bank account balance,
which is expected to disallows any activity on the account if the
balance is ever known to be less than zero.  From the perspective of
the library designer and its clients, there is no meaningful scenario
in which versions 3 and 0 can emerge from -1, since the only
transition allowed by the semantics from -1 is to itself.  Clearly,
\emph{ad hoc} mechanisms like this are error-prone and difficult to
apply in general.

@@

\begin{theorem} [\bfseries Unique LCA]
Every pair of branches in a legal history $H$, if mergeable as per
Def.~\ref{fig:mergeability}, have a unique least common ancestor.
\end{theorem}

>>>>

The problem of multiple LCAs also arises in the context of source
control systems, which employ \emph{ad hoc} mechanisms to pave the way
for three-way merging.  Git~\cite{git}, for instance,
recursively merges LCAs by default to compute a virtual ancestor, which then
serves as the LCA for merging concurrent versions. This method is
demonstrated in the branching structure for a mergeable, replicated
counter as shown in Fig.~\ref{fig:criss-cross-lcas}, where LCAs 5 and
4 of 0 and 3 are merged (with their LCA being 10) to generate -1 as
the virtual LCA to merge 5 and 4. A major downside with this approach
is that it makes no guarantees on the relationship between the virtual
ancestor and its concurrent versions; the former may not even be a
legal ancestor of the latter as per the semantics of the data type.
For instance, suppose the integer type in
Fig.~\ref{fig:criss-cross-lcas} represents a bank account balance,
which is expected to disallows any activity on the account if the
balance is ever known to be less than zero.  From the perspective of
the library designer and its clients, there is no meaningful scenario
in which versions 3 and 0 can emerge from -1, since the only
transition allowed by the semantics from -1 is to itself.  Clearly,
\emph{ad hoc} mechanisms like this are error-prone and difficult to
apply in general.

@@

\begin{theorem} [\bfseries Unique LCA]
Every pair of branches in a legal history $H$, if mergeable as per
Def.~\ref{def:mergeability}, have a unique least common ancestor.
\end{theorem}
